---
title: "Analysis"
author: "Cam Smithers"
date: "`r Sys.Date()`"
output: pdf_document
---

# Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Packages

```{r message=FALSE, warning=FALSE}
set.seed(157)
library(tidyverse)
library(glue)
library(zoo)
library(psych)
library(ggridges)
library(viridis)
library(hrbrthemes)
library(nnet)
library(ROCR)
library(vip)
library(caret)
library(rsample)
library(rstatix)
```

### Functions Script

```{r warning=FALSE}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA')
source('inR/Functions.R')
```

------------------------------------------------------------------------

## Analysis Cleaning

### Data Sets

Loading in the data sets that I'll need for analysis.

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA')
team_box_scores <- readRDS(
    'Data-NBA/Current/team_box_scores_current.rds')
player_box_scores <- readRDS(
    'Data-NBA/Current/full_box_player_current.rds')
team_season_stats <- readRDS(
    'Data-NBA/Current/team_season_stats_current.rds')
player_season_stats <- readRDS(
    'Data-NBA/Current/player_season_stats_current.rds')
```

Applying a function I wrote that fixes some aesthetic issues with player names. This will occur with foreign players that have special characters or foreign characters in their name.

```{r}
player_box_scores <- fixplayername(player_box_scores)
player_season_stats <- fixplayername(player_season_stats)
```

### Outcomes and Small Metrics

#### Outcome

I'm adding the points from the opposing team so I can create an outcome variable based on the margin of the outcome. This will be multinomial regression with four different classifications. Those are...

```{r}
opponent_stats <- team_box_scores %>%
    select(gamedate, team, pts) %>%
    rename('opp_pts'='pts')

team_box_scores <- team_box_scores %>%
    left_join(opponent_stats, by = c('gamedate', 'opponent'='team'))
```

```{r}
team_box_scores <- team_box_scores %>%
    mutate(
        ptsdiff = pts - opp_pts,
        twoleveloc = if_else(offrating > defrating, 1, 0),
        threeleveloc = case_when(
            ptsdiff >= 11 ~ 2,
            ptsdiff %in% 1:10 ~ 1,
            TRUE ~ 0),
        fourleveloc = case_when(
            ptsdiff <= -11 ~ 0,
            ptsdiff %in% -1:-10 ~ 1,
            ptsdiff %in% 1:10 ~ 2,
            ptsdiff >= 11 ~ 3)) %>%
    mutate(
        twoleveloc = factor(twoleveloc,
                            levels = c(0, 1),
                            labels = c('Loss', 'Win')),
        threeleveloc = factor(threeleveloc,
                              levels = c(0, 1, 2),
                              labels = c('Loss', 'Basic Win', 'Blowout Win')),
        fourleveloc = factor(fourleveloc,
                              levels = c(0, 1, 2, 3),
                              labels = c('Blowout Loss', 'Basic Loss', 
                                         'Basic Win', 'Blowout Win')))
```

#### Conditional Outcomes

I want to look at how missing a player affects the chance of a team to win a game; however, I have to determine which players are important for the outcome of the game. I'm going to look at few metrics and use those to determine a player's significance. For the different metrics, different players may end up being the most important player. Those metrics include...

1.  Win Shares: `ws`
2.  Value Over Replacement Player (VORP): `vorp`
3.  Player Efficiency Rating: `player_effrtg`

I'm also going to create a variable that has a unique id for each best player by a given metric.

```{r}
bp_by_ws <- best_player_metric(player_season_stats, ws, 20, 10) %>%
    rename('ws_name'='name') %>%
    unite('ws_team_id', c('year', 'team', 'ws_player_id'),
          sep = '_', remove = FALSE) %>%
    mutate(ws_team_id = paste0('ws_', ws_team_id)) %>%
    relocate(ws_team_id, .after = last_col())
bp_by_vorp <- best_player_metric(player_season_stats, vorp, 20, 10) %>%
    rename('vorp_name'='name') %>%
    unite('vorp_team_id', c('year', 'team', 'vorp_player_id'), 
          sep = '_', remove = FALSE) %>%
    mutate(vorp_team_id = paste0('vorp_', vorp_team_id)) %>%
    relocate(vorp_team_id, .after = last_col())
bp_by_per <- best_player_metric(player_season_stats, player_effrtg, 20, 10) %>%
    rename('per_name'='name') %>%
    unite('per_team_id', c('year', 'team', 'player_effrtg_player_id'), 
          sep = '_', remove = FALSE) %>%
    mutate(per_team_id = paste0('per_', per_team_id)) %>%
    relocate(per_team_id, .after = last_col())
```

Here I'm going to create a variable within each data set that will allow to check for player availability when coding a future outcome variable

Joining the data from the best player by a certain metric data sets to later determine if player was on the court and how that affected the outcome.

```{r}
team_box_scores_2 <- team_box_scores %>%
    left_join(bp_by_ws, by = c('team', 'season'='year')) %>%
    left_join(bp_by_vorp, by = c('team', 'season'='year')) %>%
    left_join(bp_by_per, by = c('team', 'season'='year')) %>%
    mutate(
        ws_tm_game_id = paste0(ws_team_id, '_', gamedate),
        vorp_tm_game_id = paste0(vorp_team_id, '_', gamedate),
        per_tm_game_id = paste0(per_team_id, '_', gamedate))

player_box_scores_2 <- player_box_scores %>%
    left_join(bp_by_ws, by = c('team', 'season'='year', 'name'='ws_name')) %>%
    left_join(bp_by_vorp, by = c('team', 'season'='year', 'name'='vorp_name')) %>%
    left_join(bp_by_per, by = c('team', 'season'='year', 'name'='per_name')) %>%
    mutate(
        ws_plyr_game_id = if_else(
            !is.na(ws_team_id), paste0(ws_team_id, '_', gamedate), NA),
        vorp_plyr_game_id = if_else(
            !is.na(vorp_team_id), paste0(vorp_team_id, '_', gamedate), NA),
        per_plyr_game_id = if_else(
            !is.na(per_team_id), paste0(per_team_id, '_', gamedate), NA))
```

After creating the data sets necessary to determine whether a team's best player (by a given metric) was available I can now create an outcome variable to check for it. I'm also going to add a variable to determines whether a game was a regular season game or playoff game.

Note: *I know I don't have to use multiple mutate functions; however, at times I like to use multiple as it helps me organize my code as I'm writing it.*

```{r}
team_box_scores_3 <- team_box_scores_2 %>%
    mutate(
        ws_bpa = if_else(ws_tm_game_id %in% 
                                player_box_scores_2$ws_plyr_game_id, 1, 0),
        vorp_bpa = if_else(vorp_tm_game_id %in% 
                               player_box_scores_2$vorp_plyr_game_id, 1, 0),
        per_bpa = if_else(per_tm_game_id %in%
                              player_box_scores_2$per_plyr_game_id, 1, 0)) %>%
    mutate(
        ws_bpa_multi = case_when(
            twoleveloc == 'Loss' & ws_bpa == 0 ~ 0,
            twoleveloc == 'Loss' & ws_bpa == 1 ~ 1,
            twoleveloc == 'Win' & ws_bpa == 0 ~ 2,
            twoleveloc == 'Win' & ws_bpa == 1 ~ 3),
        vorp_bpa_multi = case_when(
            twoleveloc == 'Loss' & vorp_bpa == 0 ~ 0,
            twoleveloc == 'Loss' & vorp_bpa == 1 ~ 1,
            twoleveloc == 'Win' & vorp_bpa == 0 ~ 2,
            twoleveloc == 'Win' & vorp_bpa == 1 ~ 3),
        per_bpa_multi = case_when(
            twoleveloc == 'Loss' & per_bpa == 0 ~ 0,
            twoleveloc == 'Loss' & per_bpa == 1 ~ 1,
            twoleveloc == 'Win' & per_bpa == 0 ~ 2,
            twoleveloc == 'Win' & per_bpa == 1 ~ 3)) %>%
    mutate(
        ws_bpa_multi = factor(ws_bpa_multi,
                              levels = c(0, 1, 2, 3),
                              labels = c('Loss Without', 'Loss With',
                                         'Win Without', 'Win With')),
        vorp_bpa_multi = factor(vorp_bpa_multi,
                                levels = c(0, 1, 2, 3),
                                labels = c('Loss Without', 'Loss With',
                                           'Win Without', 'Win With')),
        per_bpa_multi = factor(per_bpa_multi,
                               levels = c(0, 1, 2, 3),
                               labels = c('Loss Without', 'Loss With',
                                          'Win Without', 'Win With'))) %>%
    mutate(
        gametype = case_when(
            #2021 Season
            (gamedate >= as.Date('2021-05-22') & 
                 gamedate <= as.Date('2021-07-20')) |
                #2022 Season
                (gamedate >= as.Date('2022-04-16') & 
                     gamedate <= as.Date('2022-06-16')) |
                #2023 Season
                (gamedate >= as.Date('2023-04-15') & 
                     gamedate <= as.Date('2023-06-12')) |
                #2024 Season
                (gamedate >= as.Date('2024-04-20') & 
                     gamedate<= as.Date('2024-06-17')) |
                #2025 Season
                (gamedate >= as.Date('2025-04-19') & 
                     gamedate <= as.Date('2025-06-2022')) |
                #2026 Season
                 (gamedate >= as.Date('2026-04-18') & 
                     gamedate <= as.Date('2026-06-13')) ~ 'Playoff Game',
            TRUE ~ 'Regular Season Game'))
```

#### Rolling Statistics

Here I'm going to create some variables that will be used as possible predictors when testing regression models. There are great variables to work with in the data set so far; however, I want to create some rolling averages so I can measure how teams are doing across seasons and within a certain window.

```{r}
regsnbox <- team_box_scores_3 %>%
    filter(gametype == 'Regular Season Game') %>%
    group_by(team) %>%
    arrange(desc(gamedate), .by_group=TRUE) %>%
    mutate(
        roll_fg_pct = rollmean(fg_pct, k=10, fill=NA, align='left'),
        roll_threefg_pct = rollmean(threefg_pct, k=10, fill=NA, align='left'),
        roll_pts = rollmean(pts, k=10, fill=NA, align='left'),
        roll_three_par = rollmean(three_par, k=10, fill=NA, align='left'),
        roll_dreb_pct = rollmean(dreb_pct, k=10, fill=NA, align='left'),
        roll_ftar = rollmean(ftar, k=10, fill=NA, align='left'),
        roll_ast_pct = rollmean(ast_pct, k=10, fill=NA, align='left'),
        roll_offrating = rollmean(offrating, k=10, fill=NA, align='left'),
        roll_defrating = rollmean(defrating, k=10, fill=NA, align='left'),
        roll_opp_pts = rollmean(opp_pts, k=10, fill=NA, align='left'),
        roll_ptsdiff = rollmean(ptsdiff, k=10, fill=NA, align='left'),
        #Predictors for Linear Regression
        roll_ast = rollmean(ast, k=10, fill=NA, align='left'),
        roll_ts_pct = rollmean(ts_pct, k=10, fill=NA, align='left'),
        roll_pf = rollmean(pf, k=10, fill=NA, align='left'),
        roll_tov = rollmean(tov, k=10, fill=NA, align='left')) %>%
    select(gamedate, team, starts_with('roll_'))
pstsnbox <- team_box_scores_3 %>%
    filter(gametype == 'Playoff Game') %>%
    group_by(team) %>%
    arrange(desc(gamedate), .by_group=TRUE) %>%
    mutate(
        roll_fg_pct = rollmean(fg_pct, k=3, fill=NA, align='left'),
        roll_threefg_pct = rollmean(threefg_pct, k=10, fill=NA, align='left'),
        roll_pts = rollmean(pts, k=3, fill=NA, align='left'),
        roll_three_par = rollmean(three_par, k=3, fill=NA, align='left'),
        roll_dreb_pct = rollmean(dreb_pct, k=3, fill=NA, align='left'),
        roll_ftar = rollmean(ftar, k=3, fill=NA, align='left'),
        roll_ast_pct = rollmean(ast_pct, k=3, fill=NA, align='left'),
        roll_offrating = rollmean(offrating, k=3, fill=NA, align='left'),
        roll_defrating = rollmean(defrating, k=3, fill=NA, align='left'),
        roll_opp_pts = rollmean(opp_pts, k=3, fill=NA, align='left'),
        roll_ptsdiff = rollmean(ptsdiff, k=3, fill=NA, align='left'),
        #Predictors for Linear Regression
        roll_ast = rollmean(ast, k=3, fill=NA, align='left'),
        roll_ts_pct = rollmean(ts_pct, k=3, fill=NA, align='left'),
        roll_pf = rollmean(pf, k=3, fill=NA, align='left'),
        roll_tov = rollmean(tov, k=10, fill=NA, align='left')) %>%
    select(gamedate, team, starts_with('roll_'))
fullsnbox <- rbind(regsnbox, pstsnbox)
```

Using the rounding function across across various columns to clean up the rolling averages.

```{r}
team_box_scores_4 <- team_box_scores_3 %>%
    left_join(fullsnbox, by = c('gamedate', 'team')) %>%
    mutate(across(c('roll_fg_pct','roll_threefg_pct','roll_three_par',
                    'roll_ftar'), round, 3)) %>%
    mutate(across(c('roll_pts','roll_dreb_pct','roll_ast_pct',
                    'roll_offrating','roll_defrating','roll_opp_pts',
                    'roll_ptsdiff'), round, 1)) %>%
    select(-c(ws_name, ws, ws_player_id, ws_team_id, vorp_name, vorp, 
              per_name, player_effrtg, player_effrtg_player_id, 
              per_team_id, ws_tm_game_id, vorp_tm_game_id, 
              per_tm_game_id, vorp_team_id, vorp_player_id))
```

### Summary Data

Adding a non-moving average for the data set, in the off chance I may want to use it for something.

```{r}
team_box_summary <- team_box_scores_4 %>%
    group_by(season, team) %>%
    summarise(
        avg_fg_pct = round(mean(fg_pct, na.rm=TRUE), 3),
        avg_3fg_pct = round(mean(threefg_pct, na.rm=TRUE), 3),
        avg_pts = round(mean(pts, na.rm=TRUE), 1),
        avg_offrating = round(mean(offrating, na.rm=TRUE), 1),
        avg_defrating = round(mean(defrating, na.rm=TRUE), 1))
```

### Data Export for Visualizations

Export data as RDS files for visualizations in R.

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current')
saveRDS(team_box_scores_4, file = 'teamvisdata.rds')
saveRDS(team_box_summary, file = 'teamboxsummary.rds')
```

Export data as CSV files for Python dashboard.

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA/dashboard')
write_csv(team_box_scores_4, file='teamvisdata.csv')
write_csv(team_box_summary, file='teamboxsummary.csv')
```

------------------------------------------------------------------------

## Data Dive

```{r warning=FALSE}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA')
teamvisdata <- readRDS('Data-NBA/Current/teamvisdata.rds')
teamboxsummary <- readRDS('Data-NBA/Current/teamboxsummary.rds')
data_years <- sort(unique(teamvisdata$season))
```

### Summary Statistics

Here I'm using the `describe` function from the psych package to generate a table with summary statistics.

1.  N: Number of Observations
2.  Mean: Average of the Obversations
3.  SD: The Standard Deviation
4.  Median: The Number in the Middle of the Observations
5.  Trimmed: Mean Calculated Without Extreme Values
6.  MAD: Mean Absolute Deviation
7.  Min: Minimum Value
8.  Max: Maximum Value
9.  Range: Difference Between Max and Min
10. Skew: Symmetry of the Data Set
11. Kurtosis: Determine the Frequency of Outliers
12. SE: Standard Error

```{r}
describe(teamvisdata)
```

### Graphs & Plots

#### Histograms

Now let's create some basic plots to get a better understanding of our data set.

1.  Distribution of Field Goal Attempts
2.  Distribution of % of Shots Assisted On

```{r}
hist1plot <- ggplot(data=teamvisdata,
                    aes(x=fga)) +
    geom_histogram(bins=20, color='black', fill='cornflowerblue') +
    labs(
        title='Distribution of Field Goal Attempts',
        x='Field Goal Attempts',
        y='Count') +
    theme_bw()
hist1plot

hist2plot <- ggplot(data=teamvisdata,
                    aes(x=ast_pct)) +
    geom_histogram(bins=30, color='black', fill='magenta3') +
    labs(
        title='Distribution of % Shots Assisted On',
        x='% of Shots Assisted On',
        y='Count') +
    theme_bw() +
    scale_x_continuous(breaks = seq(from = 0, to = 100, by = 10))
hist2plot
```

#### Boxplots

Let's look at true shooting percentage based on the outcome of the game (win or loss with magnitude). Let's also facet it by year so we can observe any year to year trends.

```{r}
box1plot <- ggplot(data=teamvisdata,
                   aes(y=ts_pct,
                       fill=fourleveloc)) +
    geom_boxplot() +
    facet_wrap(~season) +
    labs(
        title='True Shooting % Faceted by Year',
        x='Outcome of Game',
        y='Points Scored',
        fill='Game Outcome') +
    theme_bw() +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
box1plot
```

#### Scatterplots

I want to look at the relationship between how well teams shoot the ball and how many points they score. I believe there would be a positive relationship between the two, but let's create a plot to double check. Let's also add a regression line with standard error.

```{r}
scat1plot <- ggplot(data=teamvisdata,
                    aes(x=pts,
                        y=efg_pct)) +
    geom_point(color='forestgreen') +
    labs(
        title='Points Scored vs. Effective Field Goal %',
        x='Points Scored',
        y='Effective Field Goal %') +
    theme_classic() +
    geom_smooth(method='lm', color='black', se=TRUE)
scat1plot
```

#### Time Plot

I want to take a look at my favorite team the Philadelphia 76ers, to see how they have performed on the defensive side of the ball over the past few years throughout their seasons.

```{r}
for (data_year in data_years) {
    time1plot <- ggplot(data=teamvisdata %>%
               filter(team=='PHI',
                      season==data_year,
                      gametype=='Regular Season Game'),
           aes(x=gamedate,
               y=roll_defrating)) +
#        geom_point() +
        geom_line(color=data_year) +
        theme_bw() +
        scale_x_date(date_breaks='1 month', date_labels='%b') +
        labs(
            title=glue('{data_year} Defensive Rating (Moving Average (n=10))'),
            x='Game Dates',
            y='Defensive Rating'
        )
    print(time1plot)
}
```

#### Lollipop Chart

Let's create a lollipop chart to look at teams average net rating for each season in the data set.

- Net Rating = Offensive Rating - Defensive Rating

  - Rating: Looks at how many points a teams scores or allows per 100 possessions. This allows to compare teams between games that may have had significantly different paces of play.

```{r}
data_years <- sort(unique(teamvisdata$season))


lol1plot <- ggplot(data=teamboxsummary %>%
                       mutate(avg_netrating = 
                                  avg_offrating - avg_defrating),
                   aes(x=team, y=avg_netrating, color=as.factor(season))) +
    geom_segment(
        aes(x=team, xend=team,
            y=0,yend=avg_netrating), color='black') +
    theme_bw() +
    geom_hline(yintercept=0) +
    theme(axis.text.x = element_text(angle = 90)) +
    geom_point() +
    facet_wrap(~season) +
    labs(
        title='Average Net Rating w/ Facet by Year',
        x='Team',
        y='Average Net Rating',
        color='Year')
print(lol1plot)
```

------------------------------------------------------------------------

## Logistic Regression Model Testing: Binary Game Outcome

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA')
teamvisdata <- readRDS('Data-NBA/Current/teamvisdata.rds')
teamvisdata_fixed <- na.omit(teamvisdata) %>%
    mutate(
        team = as.factor(team),
        opponent = as.factor(opponent),
        season = as.factor(season))
```

```{r}
winbinary_split <- initial_split(teamvisdata_fixed, prop=.75, strata='twoleveloc')
winbinary_train <- training(winbinary_split)
winbinary_test <- testing(winbinary_split)
```

For the models I started by adding all the variables that I wanted and deciding which ones I wanted to remove to see how the accuracy changes.

```{r}
winbinary_1 <- train(
    twoleveloc ~ roll_offrating + roll_defrating + team + homeaway + 
        vorp_bpa + roll_three_par + roll_ftar + roll_ast_pct,
    data=teamvisdata_fixed,
    method='glm',
    family='binomial',
    trControl=trainControl(method='cv', number=10))
winbinary_2 <- train(
    twoleveloc ~ roll_offrating + roll_defrating,
    data=teamvisdata_fixed,
    method='glm',
    family='binomial',
    trControl=trainControl(method='cv', number=10))
winbinary_3 <- train(
    twoleveloc ~ roll_offrating + opponent,
    data=teamvisdata_fixed,
    method='glm',
    family='binomial',
    trControl=trainControl(method='cv', number=10))
winbinary_4 <- train(
    twoleveloc ~ homeaway,
    data=teamvisdata_fixed,
    method='glm',
    family='binomial',
    trControl=trainControl(method='cv', number=10))
```

This will give the me summary statistics of the models' accuracy.

```{r}
summary(
    resamples(
        list(
            model1=winbinary_1,
            model2=winbinary_2,
            model3=winbinary_3,
            model4=winbinary_4
        )
    )
)$statistics$Accuracy
```

### ROC Curve for Logistic Regression Models

Creating a plot to show the False Positive vs. True Positive Accuracy.

```{r}
model1_prob <- predict(winbinary_1, winbinary_train, type='prob')$Win
perf1 <- prediction(model1_prob, winbinary_train$twoleveloc) %>%
    performance(measure='tpr', x.measure='fpr')

model2_prob <- predict(winbinary_2, winbinary_train, type='prob')$Win
perf2 <- prediction(model2_prob, winbinary_train$twoleveloc) %>%
    performance(measure='tpr', x.measure='fpr')

model3_prob <- predict(winbinary_3, winbinary_train, type='prob')$Win
perf3 <- prediction(model3_prob, winbinary_train$twoleveloc) %>%
    performance(measure='tpr', x.measure='fpr')

model4_prob <- predict(winbinary_4, winbinary_train, type='prob')$Win
perf4 <- prediction(model4_prob, winbinary_train$twoleveloc) %>%
    performance(measure='tpr', x.measure='fpr')

plot(perf1, col='cornflowerblue', lty=1)
plot(perf2, add=TRUE, col='magenta2', lty=2)
plot(perf3, add=TRUE, col='gold2', lty=2)
plot(perf4, add=TRUE, col='forestgreen', lty=2)
legend(0.6, 0.3, legend=c('Full Model',
                          'Net Rating',
                          'Offensive Rating w/ Opponent',
                          'Defensive Rating w/ Team'),
       col=c('cornflowerblue', 'magenta2', 'gold2', 'forestgreen'),
       lty=2:1, cex=0.6)
```

### Post-Model Training

After using machine learning to test models and determine which one performs the best, I'm going to look a bit further into the predictions that the model got wrong to better improve the model.

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA')
teamvisdata <- readRDS('Data-NBA/Current/teamvisdata.rds')
teamvisdata_fixed <- na.omit(teamvisdata) %>%
    mutate(
        team = as.factor(team),
        opponent = as.factor(opponent),
        season = as.factor(season))

bestbinary_win <- glm(twoleveloc ~ roll_offrating + roll_defrating + team + homeaway + 
                          vorp_bpa + roll_three_par + roll_ftar + roll_ast_pct,
                      data=teamvisdata_fixed, family='binomial')
bestbinary_fitted <- fitted(bestbinary_win)
bestbinary_predict <- predict(bestbinary_win, newdata=teamvisdata_fixed, type='response')
teamvisdata_fixed <- cbind(teamvisdata_fixed, bestbinary_predict)
teamvisdata_fixed <- teamvisdata_fixed %>%
    mutate(
        bestbinary_predict = round(bestbinary_predict),
           bestbinary_predict = factor(
               bestbinary_predict,
               levels = c(0, 1),
               labels = c('Loss', 'Win'))) %>%
    mutate(
        modelcheck = case_when(
            twoleveloc == bestbinary_predict ~ 'Correct',
            twoleveloc != bestbinary_predict ~ 'Incorrect',
            TRUE ~ NA))
```

#### Statistics to Investigate

I'm going to use T-test to better understand which variables are causing the model to have issues.

- **Assists: 8.06 \* 10^-5^**

- Turnover %: 3 \* 10-2

- **True Shooting %: 1.14 \* 10^-4^**

- 3PT Attempts: 2 \* 10^-2^

- **Personal Fouls: 4.41 \* 10^-3^**

- Offensive Rating: 1 \* 10^-2^

- Field Goal %:1.36 \* 10^-3^

```{r}
t.test(ts_pct ~ modelcheck, data=teamvisdata_fixed)
```

### Predicting the Unpredictable

Now that we are aware of some of the statistics that our model is having issues with, let's think how we can identify games that it would have problems with.

```{r}
#Assist Model
ast_model <- lm(ast ~ roll_ast, data=teamvisdata_fixed)
ast_fitted <- fitted(ast_model)
ast_predict <- predict(ast_model, newdata=teamvisdata_fixed)
teamvisdata_fixed <- cbind(teamvisdata_fixed, ast_predict)

#True Shooting %
ts_pct_model <- lm(ts_pct ~ roll_ts_pct, data=teamvisdata_fixed)
ts_pct_fitted <- fitted(ts_pct_model)
ts_pct_predict <- predict(ts_pct_model, newdata=teamvisdata_fixed)
teamvisdata_fixed <- cbind(teamvisdata_fixed, ts_pct_predict)

#Personal Fouls
pf_model <- lm(pf ~ roll_pf, data=teamvisdata_fixed)
pf_fitted <- fitted(pf_model)
pf_predict <- predict(pf_model, newdata=teamvisdata_fixed)
teamvisdata_fixed <- cbind(teamvisdata_fixed, pf_predict)

teamvisdata_fixed <- teamvisdata_fixed %>%
    mutate(
        ast_predict = round(ast_predict, 1),
        ts_pct_predict = round(ts_pct_predict, 3),
        pf_predict = round(pf_predict, 1))

teamvisdata_fixed <- na.omit(teamvisdata_fixed)

binarycheck_split <- initial_split(teamvisdata_fixed, prop=.75, strata='modelcheck')
binarycheck_train <- training(binarycheck_split)
binarycheck_test <- testing(binarycheck_split)

binarycheck1 <- train(
    modelcheck ~ .,
    data=teamvisdata_fixed,
    method='glm',
    family='binomial',
    trControl=trainControl(method='cv', number=10))

vip(binarycheck1, num_features=15)
```

Above we have a graph that displays the top 15 influential variables when trying to estimate when the original model will be wrong. Of these variables many of them are directly related to the outcome of the game (which we are trying to predict) or we won't know until the game is over, so we can ignore those. I would like to focus on `per_bpa` and `vorp_bpa`, these variables indicate whether the best player on a team (by a specific metric) will not play during that game. If we wanted to use this model to pick the winner of a game where a team is missing their player with the highest PER (Player Efficiency Rating) or VORP (Value Over Replacement Player), we have two options.

1.  We can pass on predicting that game entirely.
    - This method allows us to focus on games where we are confident in our predictions and not worry about where our parlay may cash.
2.  Determine how each team performs without that specific player, then adjust our prediction from there.
    - This method would require us to develop more statistical models, specifically those that are focused on a specific team in a specific situation. I few this method as advantageous only if you are **only** focused on the outcomes of a specific team.
