---
title: "Analysis"
author: "Cam Smithers"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    fig_height: 5
---

# Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Packages

```{r}
library(tidyverse)
```

### Functions Script

```{r}
source('/Users/camsmithers/Desktop/Camalytics/NBA/inR/Functions.R')
```

### Data Sets

Loading in the data sets that I'll need for analysis.

```{r}
team_box_scores <- readRDS(
    '/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current/team_box_scores_updated.rds')
player_box_scores <- readRDS(
    '/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current/full_box_player_updated.rds')
team_season_stats <- readRDS(
    '/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current/team_season_stats_updated.rds')
player_season_stats <- readRDS(
    '/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current/player_season_stats_updated.rds')
```

Applying a function I wrote that fixes some aesthetic issues with player names. This will occur with foreign players that have special characters or foreign characters in their name.

```{r}
player_box_scores <- fixplayername(player_box_scores)
player_season_stats <- fixplayername(player_season_stats)
```

### Outcomes and Small Metrics

#### Outcome

I'm adding the points from the opposing team so I can create an outcome variable based on the margin of the outcome. This will be multinomial regression with four different classifications. Those are...

```{r}
opponent_stats <- team_box_scores %>%
    select(gamedate, team, pts) %>%
    rename('opp_pts'='pts')

team_box_scores <- team_box_scores %>%
    left_join(opponent_stats, by = c('gamedate', 'opponent'='team'))
```

```{r}
team_box_scores <- team_box_scores %>%
    mutate(
        ptsdiff = pts - opp_pts,
        twoleveloc = ifelse(offrating > defrating, 1, 0),
        threeleveloc = case_when(
            ptsdiff >= 11 ~ 2,
            ptsdiff %in% 1:10 ~ 1,
            TRUE ~ 0),
        fourleveloc = case_when(
            ptsdiff <= -11 ~ 0,
            ptsdiff %in% -1:-10 ~ 1,
            ptsdiff %in% 1:10 ~ 2,
            ptsdiff >= 11 ~ 3)) %>%
    mutate(
        twoleveloc = factor(twoleveloc,
                            levels = c(0, 1),
                            labels = c('Loss', 'Win')),
        threeleveloc = factor(threeleveloc,
                              levels = c(0, 1, 2),
                              labels = c('Loss', 'Basic Win', 'Blowout Win')),
        fourleveloc = factor(fourleveloc,
                              levels = c(0, 1, 2, 3),
                              labels = c('Blowout Loss', 'Basic Loss', 'Basic Win', 'Blowout Win')))
```

### Outcome Visualizations

#### Plot 1: Wins vs. Losses

```{r, fig.align='left'}
plot1 <- ggplot(team_box_scores,
                aes(x=twoleveloc,
                    fill=twoleveloc)) + 
    geom_bar(position='dodge', color='black') + 
    labs(
        title='Wins vs. Losses from 2021 to 2025',
        x='Win & Loss Count',
        y='Count',
        fill='Win vs. Loss') +
    facet_wrap(~team) + 
    theme_bw()
plot1
```

#### Plot 2: 4 Level Outcome Variable

```{r, fig.align='left'}
plot2 <- ggplot(team_box_scores,
                aes(x=fourleveloc,
                    fill=fourleveloc)) + 
    geom_bar(position='dodge', color='black') + 
    labs(
        title='Magnitude of Wins vs. Losses from 2021 to 2025',
        x='Win & Loss Count',
        y='Count',
        fill='Win vs. Loss Magnitude') +
    facet_wrap(~team) + 
    theme_bw()
plot2
```

I want to look at how missing a player affects the chance of a team to win a game; however, I have to determine which players are important for the outcome of the game. I'm going to look at few metrics and use those to determine a player's significance. For the different metrics, different players may end up being the most important player. Those metrics include...

1.  Win Shares: `ws`
2.  Value Over Replacement Player (VORP): `vorp`
3.  Player Efficiency Rating: `player_effrtg`

```{r}
bp_by_ws <- best_player_metric(player_season_stats, ws, 20, 10)
bp_by_vorp <- best_player_metric(player_season_stats, vorp, 20, 10)
bp_by_per <- best_player_metric(player_season_stats, player_effrtg, 20, 10)
```

Joining the data from the best player by a certain metric data sets to later determine if player was on the court and how that affected the outcome.

```{r}
team_box_scores_2 <- team_box_scores %>%
    left_join(bp_by_ws, by = c('team', 'season'='year')) %>%
    left_join(bp_by_vorp, by = c('team', 'season'='year', 'name')) %>%
    left_join(bp_by_per, by = c('team', 'season'='year', 'name'))
```

Things needed to create this conditional outcome variable or variables...
