---
title: "Analysis"
author: "Cam Smithers"
date: "`r Sys.Date()`"
output: pdf_document
---

# Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Packages

```{r message=FALSE, warning=FALSE}
set.seed(157)
library(tidyverse)
library(glue)
library(zoo)
library(psych)
```

### Functions Script

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA')
source('inR/Functions.R')
```

## Analysis Cleaning

### Data Sets

Loading in the data sets that I'll need for analysis.

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA')
team_box_scores <- readRDS(
    'Data-NBA/Current/team_box_scores_current.rds')
player_box_scores <- readRDS(
    'Data-NBA/Current/full_box_player_current.rds')
team_season_stats <- readRDS(
    'Data-NBA/Current/team_season_stats_current.rds')
player_season_stats <- readRDS(
    'Data-NBA/Current/player_season_stats_current.rds')
```

Applying a function I wrote that fixes some aesthetic issues with player names. This will occur with foreign players that have special characters or foreign characters in their name.

```{r}
player_box_scores <- fixplayername(player_box_scores)
player_season_stats <- fixplayername(player_season_stats)
```

### Outcomes and Small Metrics

#### Outcome

I'm adding the points from the opposing team so I can create an outcome variable based on the margin of the outcome. This will be multinomial regression with four different classifications. Those are...

```{r}
opponent_stats <- team_box_scores %>%
    select(gamedate, team, pts) %>%
    rename('opp_pts'='pts')

team_box_scores <- team_box_scores %>%
    left_join(opponent_stats, by = c('gamedate', 'opponent'='team'))
```

```{r}
team_box_scores <- team_box_scores %>%
    mutate(
        ptsdiff = pts - opp_pts,
        twoleveloc = if_else(offrating > defrating, 1, 0),
        threeleveloc = case_when(
            ptsdiff >= 11 ~ 2,
            ptsdiff %in% 1:10 ~ 1,
            TRUE ~ 0),
        fourleveloc = case_when(
            ptsdiff <= -11 ~ 0,
            ptsdiff %in% -1:-10 ~ 1,
            ptsdiff %in% 1:10 ~ 2,
            ptsdiff >= 11 ~ 3)) %>%
    mutate(
        twoleveloc = factor(twoleveloc,
                            levels = c(0, 1),
                            labels = c('Loss', 'Win')),
        threeleveloc = factor(threeleveloc,
                              levels = c(0, 1, 2),
                              labels = c('Loss', 'Basic Win', 'Blowout Win')),
        fourleveloc = factor(fourleveloc,
                              levels = c(0, 1, 2, 3),
                              labels = c('Blowout Loss', 'Basic Loss', 
                                         'Basic Win', 'Blowout Win')))
```

#### Conditional Outcomes

I want to look at how missing a player affects the chance of a team to win a game; however, I have to determine which players are important for the outcome of the game. I'm going to look at few metrics and use those to determine a player's significance. For the different metrics, different players may end up being the most important player. Those metrics include...

1.  Win Shares: `ws`
2.  Value Over Replacement Player (VORP): `vorp`
3.  Player Efficiency Rating: `player_effrtg`

I'm also going to create a variable that has a unique id for each best player by a given metric.

```{r}
bp_by_ws <- best_player_metric(player_season_stats, ws, 20, 10) %>%
    rename('ws_name'='name') %>%
    unite('ws_team_id', c('year', 'team', 'ws_player_id'),
          sep = '_', remove = FALSE) %>%
    mutate(ws_team_id = paste0('ws_', ws_team_id)) %>%
    relocate(ws_team_id, .after = last_col())
bp_by_vorp <- best_player_metric(player_season_stats, vorp, 20, 10) %>%
    rename('vorp_name'='name') %>%
    unite('vorp_team_id', c('year', 'team', 'vorp_player_id'), 
          sep = '_', remove = FALSE) %>%
    mutate(vorp_team_id = paste0('vorp_', vorp_team_id)) %>%
    relocate(vorp_team_id, .after = last_col())
bp_by_per <- best_player_metric(player_season_stats, player_effrtg, 20, 10) %>%
    rename('per_name'='name') %>%
    unite('per_team_id', c('year', 'team', 'player_effrtg_player_id'), 
          sep = '_', remove = FALSE) %>%
    mutate(per_team_id = paste0('per_', per_team_id)) %>%
    relocate(per_team_id, .after = last_col())
```

Here I'm going to create a variable within each data set that will allow to check for player availability when coding a future outcome variable

Joining the data from the best player by a certain metric data sets to later determine if player was on the court and how that affected the outcome.

```{r}
team_box_scores_2 <- team_box_scores %>%
    left_join(bp_by_ws, by = c('team', 'season'='year')) %>%
    left_join(bp_by_vorp, by = c('team', 'season'='year')) %>%
    left_join(bp_by_per, by = c('team', 'season'='year')) %>%
    mutate(
        ws_tm_game_id = paste0(ws_team_id, '_', gamedate),
        vorp_tm_game_id = paste0(vorp_team_id, '_', gamedate),
        per_tm_game_id = paste0(per_team_id, '_', gamedate))

player_box_scores_2 <- player_box_scores %>%
    left_join(bp_by_ws, by = c('team', 'season'='year', 'name'='ws_name')) %>%
    left_join(bp_by_vorp, by = c('team', 'season'='year', 'name'='vorp_name')) %>%
    left_join(bp_by_per, by = c('team', 'season'='year', 'name'='per_name')) %>%
    mutate(
        ws_plyr_game_id = if_else(
            !is.na(ws_team_id), paste0(ws_team_id, '_', gamedate), NA),
        vorp_plyr_game_id = if_else(
            !is.na(vorp_team_id), paste0(vorp_team_id, '_', gamedate), NA),
        per_plyr_game_id = if_else(
            !is.na(per_team_id), paste0(per_team_id, '_', gamedate), NA))
```

After creating the data sets necessary to determine whether a team's best player (by a given metric) was available I can now create an outcome variable to check for it. I'm also going to add a variable to determines whether a game was a regular season game or playoff game.

Note: *I know I don't have to use multiple mutate functions; however, at times I like to use multiple as it helps me organize my code as I'm writing it.*

```{r}
team_box_scores_3 <- team_box_scores_2 %>%
    mutate(
        ws_bpa = if_else(ws_tm_game_id %in% 
                                player_box_scores_2$ws_plyr_game_id, 1, 0),
        vorp_bpa = if_else(vorp_tm_game_id %in% 
                               player_box_scores_2$vorp_plyr_game_id, 1, 0),
        per_bpa = if_else(per_tm_game_id %in%
                              player_box_scores_2$per_plyr_game_id, 1, 0)) %>%
    mutate(
        ws_bpa_multi = case_when(
            twoleveloc == 'Loss' & ws_bpa == 0 ~ 0,
            twoleveloc == 'Loss' & ws_bpa == 1 ~ 1,
            twoleveloc == 'Win' & ws_bpa == 0 ~ 2,
            twoleveloc == 'Win' & ws_bpa == 1 ~ 3),
        vorp_bpa_multi = case_when(
            twoleveloc == 'Loss' & vorp_bpa == 0 ~ 0,
            twoleveloc == 'Loss' & vorp_bpa == 1 ~ 1,
            twoleveloc == 'Win' & vorp_bpa == 0 ~ 2,
            twoleveloc == 'Win' & vorp_bpa == 1 ~ 3),
        per_bpa_multi = case_when(
            twoleveloc == 'Loss' & per_bpa == 0 ~ 0,
            twoleveloc == 'Loss' & per_bpa == 1 ~ 1,
            twoleveloc == 'Win' & per_bpa == 0 ~ 2,
            twoleveloc == 'Win' & per_bpa == 1 ~ 3)) %>%
    mutate(
        ws_bpa_multi = factor(ws_bpa_multi,
                              levels = c(0, 1, 2, 3),
                              labels = c('Loss Without', 'Loss With',
                                         'Win Without', 'Win With')),
        vorp_bpa_multi = factor(vorp_bpa_multi,
                                levels = c(0, 1, 2, 3),
                                labels = c('Loss Without', 'Loss With',
                                           'Win Without', 'Win With')),
        per_bpa_multi = factor(per_bpa_multi,
                               levels = c(0, 1, 2, 3),
                               labels = c('Loss Without', 'Loss With',
                                          'Win Without', 'Win With'))) %>%
    mutate(
        gametype = case_when(
            #2021 Season
            (gamedate >= as.Date('2021-05-22') & 
                 gamedate <= as.Date('2021-07-20')) |
                #2022 Season
                (gamedate >= as.Date('2022-04-16') & 
                     gamedate <= as.Date('2022-06-16')) |
                #2023 Season
                (gamedate >= as.Date('2023-04-15') & 
                     gamedate <= as.Date('2023-06-12')) |
                #2024 Season
                (gamedate >= as.Date('2024-04-20') & 
                     gamedate<= as.Date('2024-06-17')) |
                #2025 Season
                (gamedate >= as.Date('2025-04-19') & 
                     gamedate <= as.Date('2025-06-2022')) |
                #2026 Season
                 (gamedate >= as.Date('2026-04-18') & 
                     gamedate <= as.Date('2026-06-13')) ~ 'Playoff Game',
            TRUE ~ 'Regular Season Game'))
```

Here I'm going to create some variables that will be used as possible predictors when testing regression models. There are great variables to work with in the data set so far; however, I want to create some rolling averages so I can measure how teams are doing across seasons and within a certain window.

```{r}
regsnbox <- team_box_scores_3 %>%
    filter(gametype == 'Regular Season Game') %>%
    group_by(team) %>%
    arrange(desc(gamedate), .by_group=TRUE) %>%
    mutate(
        roll_fg_pct = rollmean(fg_pct, k=10, fill=NA, align='left'),
        roll_threefg_pct = rollmean(threefg_pct, k=10, fill=NA, align='left'),
        roll_pts = rollmean(pts, k=10, fill=NA, align='left'),
        roll_three_par = rollmean(three_par, k=10, fill=NA, align='left'),
        roll_dreb_pct = rollmean(dreb_pct, k=10, fill=NA, align='left'),
        roll_ftar = rollmean(ftar, k=10, fill=NA, align='left'),
        roll_ast_pct = rollmean(ast_pct, k=10, fill=NA, align='left'),
        roll_offrating = rollmean(offrating, k=10, fill=NA, align='left'),
        roll_defrating = rollmean(defrating, k=10, fill=NA, align='left'),
        roll_opp_pts = rollmean(opp_pts, k=10, fill=NA, align='left'),
        roll_ptsdiff = rollmean(ptsdiff, k=10, fill=NA, align='left')) %>%
    select(gamedate, team, starts_with('roll_'))
pstsnbox <- team_box_scores_3 %>%
    filter(gametype == 'Playoff Game') %>%
    group_by(team) %>%
    arrange(desc(gamedate), .by_group=TRUE) %>%
    mutate(
        roll_fg_pct = rollmean(fg_pct, k=3, fill=NA, align='left'),
        roll_threefg_pct = rollmean(threefg_pct, k=10, fill=NA, align='left'),
        roll_pts = rollmean(pts, k=3, fill=NA, align='left'),
        roll_three_par = rollmean(three_par, k=3, fill=NA, align='left'),
        roll_dreb_pct = rollmean(dreb_pct, k=3, fill=NA, align='left'),
        roll_ftar = rollmean(ftar, k=3, fill=NA, align='left'),
        roll_ast_pct = rollmean(ast_pct, k=3, fill=NA, align='left'),
        roll_offrating = rollmean(offrating, k=3, fill=NA, align='left'),
        roll_defrating = rollmean(defrating, k=3, fill=NA, align='left'),
        roll_opp_pts = rollmean(opp_pts, k=3, fill=NA, align='left'),
        roll_ptsdiff = rollmean(ptsdiff, k=3, fill=NA, align='left')) %>%
    select(gamedate, team, starts_with('roll_'))
fullsnbox <- rbind(regsnbox, pstsnbox)
```

Using the rounding function across across various columns to clean up the rolling averages.

```{r}
team_box_scores_4 <- team_box_scores_3 %>%
    left_join(fullsnbox, by = c('gamedate', 'team')) %>%
    mutate(across(c('roll_fg_pct','roll_threefg_pct','roll_three_par',
                    'roll_ftar'), round, 3)) %>%
    mutate(across(c('roll_pts','roll_dreb_pct','roll_ast_pct',
                    'roll_offrating','roll_defrating','roll_opp_pts',
                    'roll_ptsdiff'), round, 1)) %>%
    select(-c(ws_name, ws, ws_player_id, ws_team_id, vorp_name, vorp, 
              per_name, player_effrtg, player_effrtg_player_id, 
              per_team_id, ws_tm_game_id, vorp_tm_game_id, 
              per_tm_game_id, vorp_team_id, vorp_player_id))
```

### Summary Data

Adding a non-moving average for the data set, in the off chance I may want to use it for something.

```{r}
team_box_summary <- team_box_scores_4 %>%
    group_by(season, team) %>%
    summarise(
        avg_fg_pct = round(mean(fg_pct, na.rm=TRUE), 3),
        avg_3fg_pct = round(mean(threefg_pct, na.rm=TRUE), 3),
        avg_pts = round(mean(pts, na.rm=TRUE), 1),
        avg_offrating = round(mean(offrating, na.rm=TRUE), 1),
        avg_defrating = round(mean(defrating, na.rm=TRUE), 1))
```

### Data Export for Visualizations

Export data as RDS files for visualizations in R.

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current')
saveRDS(team_box_scores_4, file = 'teamvisdata.rds')
saveRDS(team_box_summary, file = 'teamboxsummary.rds')
```

Export data as CSV files for Python dashboard.

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA/dashboard')
write_csv(team_box_scores_4, file='teamvisdata.csv')
write_csv(team_box_summary, file='teamboxsummary.csv')
```

------------------------------------------------------------------------

## Basic Data Dive

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA')
teamvisdata <- readRDS('Data-NBA/Current/teamvisdata.rds')
```

Here I'm using the `describe` function from the psych package to generate a table with summary statistics.

1.  N: Number of Observations
2.  Mean: Average of the Obversations
3.  SD: The Standard Deviation
4.  Median: The Number in the Middle of the Observations
5.  Trimmed: Mean Calculated Without Extreme Values
6.  MAD: Mean Absolute Deviation
7.  Min: Minimum Value
8.  Max: Maximum Value
9.  Range: Difference Between Max and Min
10. Skew: Symmetry of the Data Set
11. Kurtosis: Determine the Frequency of Outliers
12. SE: Standard Error

```{r}
describe(teamvisdata)
```

Now let's create some basic plots to get a better understanding of our data set.

```{r}
hist1plot <- ggplot(data=teamvisdata,
                    aes(x=fga)) +
    geom_histogram(bins=30, color='black', fill='cornflowerblue') +
    labs(
        title='Distribution of Field Goal Attempts',
        x='Field Goal Attempts',
        y='Count') +
    theme_bw() +
    scale_x_continuous(breaks = seq(from = 60, to = 130, by = 5))
hist1plot

hist2plot <- ggplot(data=teamvisdata,
                    aes(x=ast_pct)) +
    geom_histogram(bins=30, color='black', fill='magenta3') +
    labs(
        title='Distribution of Shots Assisted On',
        x='% of Shots Assisted On',
        y='Count') +
    theme_bw() +
    scale_x_continuous(breaks = seq(from = 0, to = 100, by = 10))
hist2plot
```

```{r}
scat1plot <- ggplot(data=teamvisdata %>%
                        filter(season==2023),
                    aes(x=ts_pct,
                        y=ftar)) +
    geom_point(color='forestgreen') +
    labs(
        title='True Shooting % vs. Free Throw Attempt Rate',
        x='True Shooting %',
        y='Free Throw Attempt Rate') +
    theme_classic()
scat1plot
```
