---
title: "Analysis"
author: "Cam Smithers"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    fig_height: 5
---

# Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = '/Users/camsmithers/Desktop/Camalytics/NBA')
```

### Packages

```{r}
set.seed(157)
library(tidyverse)
```

### Functions Script

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA')
source('inR/Functions.R')
```

### Data Sets

Loading in the data sets that I'll need for analysis.

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA')
team_box_scores <- readRDS(
    'Data-NBA/Current/team_box_scores_updated.rds')
player_box_scores <- readRDS(
    'Data-NBA/Current/full_box_player_updated.rds')
team_season_stats <- readRDS(
    'Data-NBA/Current/team_season_stats_updated.rds')
player_season_stats <- readRDS(
    'Data-NBA/Current/player_season_stats_updated.rds')
```

Applying a function I wrote that fixes some aesthetic issues with player names. This will occur with foreign players that have special characters or foreign characters in their name.

```{r}
player_box_scores <- fixplayername(player_box_scores)
player_season_stats <- fixplayername(player_season_stats)
```

### Outcomes and Small Metrics

#### Outcome

I'm adding the points from the opposing team so I can create an outcome variable based on the margin of the outcome. This will be multinomial regression with four different classifications. Those are...

```{r}
opponent_stats <- team_box_scores %>%
    select(gamedate, team, pts) %>%
    rename('opp_pts'='pts')

team_box_scores <- team_box_scores %>%
    left_join(opponent_stats, by = c('gamedate', 'opponent'='team'))
```

```{r}
team_box_scores <- team_box_scores %>%
    mutate(
        ptsdiff = pts - opp_pts,
        twoleveloc = if_else(offrating > defrating, 1, 0),
        threeleveloc = case_when(
            ptsdiff >= 11 ~ 2,
            ptsdiff %in% 1:10 ~ 1,
            TRUE ~ 0),
        fourleveloc = case_when(
            ptsdiff <= -11 ~ 0,
            ptsdiff %in% -1:-10 ~ 1,
            ptsdiff %in% 1:10 ~ 2,
            ptsdiff >= 11 ~ 3)) %>%
    mutate(
        twoleveloc = factor(twoleveloc,
                            levels = c(0, 1),
                            labels = c('Loss', 'Win')),
        threeleveloc = factor(threeleveloc,
                              levels = c(0, 1, 2),
                              labels = c('Loss', 'Basic Win', 'Blowout Win')),
        fourleveloc = factor(fourleveloc,
                              levels = c(0, 1, 2, 3),
                              labels = c('Blowout Loss', 'Basic Loss', 
                                         'Basic Win', 'Blowout Win')))
```

#### Conditional Outcomes

I want to look at how missing a player affects the chance of a team to win a game; however, I have to determine which players are important for the outcome of the game. I'm going to look at few metrics and use those to determine a player's significance. For the different metrics, different players may end up being the most important player. Those metrics include...

1.  Win Shares: `ws`
2.  Value Over Replacement Player (VORP): `vorp`
3.  Player Efficiency Rating: `player_effrtg`

I'm also going to create a variable that has a unique id for each best player by a given metric.

```{r}
bp_by_ws <- best_player_metric(player_season_stats, ws, 20, 10) %>%
    rename('ws_name'='name') %>%
    unite('ws_team_id', c('year', 'team', 'ws_player_id'),
          sep = '_', remove = FALSE) %>%
    mutate(ws_team_id = paste0('ws_', ws_team_id)) %>%
    relocate(ws_team_id, .after = last_col())
bp_by_vorp <- best_player_metric(player_season_stats, vorp, 20, 10) %>%
    rename('vorp_name'='name') %>%
    unite('vorp_team_id', c('year', 'team', 'vorp_player_id'), 
          sep = '_', remove = FALSE) %>%
    mutate(vorp_team_id = paste0('vorp_', vorp_team_id)) %>%
    relocate(vorp_team_id, .after = last_col())
bp_by_per <- best_player_metric(player_season_stats, player_effrtg, 20, 10) %>%
    rename('per_name'='name') %>%
    unite('per_team_id', c('year', 'team', 'player_effrtg_player_id'), 
          sep = '_', remove = FALSE) %>%
    mutate(per_team_id = paste0('per_', per_team_id)) %>%
    relocate(per_team_id, .after = last_col())
```

Here I'm going to create a variable within each data set that will allow to check for player availability when coding a future outcome variable

Joining the data from the best player by a certain metric data sets to later determine if player was on the court and how that affected the outcome.

```{r}
team_box_scores_2 <- team_box_scores %>%
    left_join(bp_by_ws, by = c('team', 'season'='year')) %>%
    left_join(bp_by_vorp, by = c('team', 'season'='year')) %>%
    left_join(bp_by_per, by = c('team', 'season'='year')) %>%
    mutate(
        ws_tm_game_id = paste0(ws_team_id, '_', gamedate),
        vorp_tm_game_id = paste0(vorp_team_id, '_', gamedate),
        per_tm_game_id = paste0(per_team_id, '_', gamedate))

player_box_scores_2 <- player_box_scores %>%
    left_join(bp_by_ws, by = c('team', 'season'='year', 'name'='ws_name')) %>%
    left_join(bp_by_vorp, by = c('team', 'season'='year', 'name'='vorp_name')) %>%
    left_join(bp_by_per, by = c('team', 'season'='year', 'name'='per_name')) %>%
    mutate(
        ws_plyr_game_id = if_else(
            !is.na(ws_team_id), paste0(ws_team_id, '_', gamedate), NA),
        vorp_plyr_game_id = if_else(
            !is.na(vorp_team_id), paste0(vorp_team_id, '_', gamedate), NA),
        per_plyr_game_id = if_else(
            !is.na(per_team_id), paste0(per_team_id, '_', gamedate), NA))
```

After creating the data sets necessary to determine whether a team's best player (by a given metric) was available I can now create an outcome variable to check for it.

Note: *I know I don't have to use multiple mutate functions; however, at times I like to use multiple as it helps me organize my code as I'm writing it.*

```{r}
team_box_scores_3 <- team_box_scores_2 %>%
    mutate(
        ws_bpa = if_else(ws_tm_game_id %in% 
                                player_box_scores_2$ws_plyr_game_id, 1, 0),
        vorp_bpa = if_else(vorp_tm_game_id %in% 
                               player_box_scores_2$vorp_plyr_game_id, 1, 0),
        per_bpa = if_else(per_tm_game_id %in%
                              player_box_scores_2$per_plyr_game_id, 1, 0)) %>%
    mutate(
        ws_bpa_multi = case_when(
            twoleveloc == 'Loss' & ws_bpa == 0 ~ 0,
            twoleveloc == 'Loss' & ws_bpa == 1 ~ 1,
            twoleveloc == 'Win' & ws_bpa == 0 ~ 2,
            twoleveloc == 'Win' & ws_bpa == 1 ~ 3),
        vorp_bpa_multi = case_when(
            twoleveloc == 'Loss' & vorp_bpa == 0 ~ 0,
            twoleveloc == 'Loss' & vorp_bpa == 1 ~ 1,
            twoleveloc == 'Win' & vorp_bpa == 0 ~ 2,
            twoleveloc == 'Win' & vorp_bpa == 1 ~ 3),
        per_bpa_multi = case_when(
            twoleveloc == 'Loss' & per_bpa == 0 ~ 0,
            twoleveloc == 'Loss' & per_bpa == 1 ~ 1,
            twoleveloc == 'Win' & per_bpa == 0 ~ 2,
            twoleveloc == 'Win' & per_bpa == 1 ~ 3)) %>%
    mutate(
        ws_bpa_multi = factor(ws_bpa_multi,
                              levels = c(0, 1, 2, 3),
                              labels = c('Loss Without', 'Loss With',
                                         'Win Without', 'Win With')),
        vorp_bpa_multi = factor(vorp_bpa_multi,
                                levels = c(0, 1, 2, 3),
                                labels = c('Loss Without', 'Loss With',
                                           'Win Without', 'Win With')),
        per_bpa_multi = factor(per_bpa_multi,
                               levels = c(0, 1, 2, 3),
                               labels = c('Loss Without', 'Loss With',
                                          'Win Without', 'Win With')))
```

### Summary Data

```{r}
team_box_summary <- team_box_scores_3 %>%
    group_by(season, team) %>%
    summarise(
        avg_fg_pct = round(mean(fg_pct, na.rm=TRUE), 3),
        avg_3fg_pct = round(mean(threefg_pct, na.rm=TRUE), 3),
        avg_pts = round(mean(pts, na.rm=TRUE), 1),
        avg_offrating = round(mean(offrating, na.rm=TRUE), 1),
        avg_defrating = round(mean(defrating, na.rm=TRUE), 1))
```

### Data Export for Visualizations

```{r}
setwd('/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current')
saveRDS(team_box_scores_3, file = 'teamvisdata.rds')
```
