---
title: "Analysis"
author: "Cam Smithers"
date: "`r Sys.Date()`"
output: pdf_document
---

# Basic Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Packages

```{r}
library(tidyverse)
```

### Functions Script

```{r}
source('/Users/camsmithers/Desktop/Camalytics/NBA/inR/Functions.R')
```

### Data Sets

Loading in the data sets that I'll need for analysis.

```{r}
team_box_scores <- readRDS(
    '/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current/team_box_scores_updated.rds')
player_box_scores <- readRDS(
    '/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current/full_box_player_updated.rds')
team_season_stats <- readRDS(
    '/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current/team_season_stats_updated.rds')
player_season_stats <- readRDS(
    '/Users/camsmithers/Desktop/Camalytics/NBA/Data-NBA/Current/player_season_stats_updated.rds')
```

Applying a function I wrote that fixes some aesthetic issues with player names. This will occur with foreign players that have special characters or foreign characters in their name.

```{r}
player_box_scores <- fixplayername(player_box_scores)
player_season_stats <- fixplayername(player_season_stats)
```

### Outcomes and Small Metrics

#### Outcome

I'm adding the points from the opposing team so I can create an outcome variable based on the margin of the outcome. This will be multinomial regression with four different classifications. Those are...

1.  Win by 11+
2.  Win by 1 to 10
3.  Lose by 1 to 10
4.  Lose by 11+

```{r}
opponent_stats <- team_box_scores %>%
    select(gamedate, team, pts) %>%
    rename('opp_pts'='pts')

team_box_scores <- team_box_scores %>%
    left_join(opponent_stats, by = c('gamedate', 'opponent'='team'))
```

```{r}
team_box_scores <- team_box_scores %>%
    mutate(
        win = ifelse(offrating > defrating, 1, 0), #game outcome (binary)
        ptsdiff = pts - opp_pts,
        threeleveloc = case_when(
            ptsdiff %in% 1:10 ~ 0,
            ptsdiff >= 11 ~ 1,
            TRUE ~ 2),
        fourlevelloc = case_when(
            ptsdiff 
        )
    )
```

I want to look at how missing a player affects the chance of a team to win a game; however, I have to determine which players are important for the outcome of the game. I'm going to look at few metrics and use those to determine a player's significance. For the different metrics, different players may end up being the most important player. Those metrics include...

1.  Win Shares: `ws`
2.  Value Over Replacement Player (VORP): `vorp`
3.  Player Efficiency Rating: `player_effrtg`

```{r}
bp_by_ws <- best_player_metric(player_season_stats, ws, 20, 10)
bp_by_vorp <- best_player_metric(player_season_stats, vorp, 20, 10)
bp_by_per <- best_player_metric(player_season_stats, player_effrtg, 20, 10)
```

Joining the data from the best player by a certain metric data sets to later determine if player was on the court and how that affected the outcome.

```{r}
team_box_scores_2 <- team_box_scores %>%
    left_join(bp_by_ws, by = c('team', 'season'='year')) %>%
    left_join(bp_by_vorp, by = c('team', 'season'='year', 'name')) %>%
    left_join(bp_by_per, by = c('team', 'season'='year', 'name'))
```
